Basic Parameter estimation of Binary Neutron Star Systems by the Advanced LIGO/Virgo Network

### 2013-12-12 doc/Review1Response.txt

1st referee

The paper presents a numerical calculation on the measurement of BNS physical parameters, from GW signals to be obtained with the advanced detector array.
The paper is clearly written and the calculation look sound and free of basic errors.
The analysis reported here is a sort of reduced version, at higher SNR, of the analysis reported in Nissanke et al, 2013, Astrophys. J., 767, 124, 124.
Both analyses use the zero-noise, MCMC technique, and their results are consistent.
I am not convinced that the paper, at this stage of the work, provides substantial new information over that of Nissanke. The only piece of information that I find new is to be found in figure 7 and 8, wherein the positional accuracy is reported as a function of the sky location of the source.
As most of methods and procedures are not new, I think the authors, after having substantially shortened the introductory parts and minimized the description of the zero-noise MCMC method, may want to highlight the delta knowledge contributed by this paper relative to the recent literature.

I have also one comment on section 3.4 and the related figure 9.

At face value figure 9 would indicate that by averaging overn many non zero-noise MCMC calculation, the distribution of all parameters would be substantially smeared out . The authors state instead that the figure demonstrates that the distribution stays the same as in the zero-noise case. I think this is confusing and should be clarified.

I think the authors have to respond to these objections before the paper may be considered for publication.

———————————————————————————————————

The reviewer brings up two points where our paper can be improved: the length of the description of our MCMC zero-noise choice, and the difference between our work and the work of Nissanke et al.  We discuss these points below:

Length of the zero-noise sections:

We feel that the length of the description of our MCMC technique is justified for two reasons:

	1.the code employed is distinct from that of Nissanke et al.  While it is certainly true that the two codes use the same technique (an MCMC is an MCMC, after all), the LALInference_MCMC code was specifically developed for the analysis of raw data from the LIGO instrument, and is the only MCMC code to have been tested against real data, and

	2. The description of our Bayesian technique is necessary, due to the difference in priors between our study and Nissanke. In particular, we employ a flat prior in the component masses (versus Nissanke’s flat prior in the chirp mass), which can lead to substantially different recovered posteriors

Delta Knowledge between our work and Nissanke et al:

The second issue is, of course, what additional information is provided by our paper that was not provided by Nissanke at al. 2011 or 2013.  There are two primary areas in which our study differentiates itself from the two of Nissanke, in addition to the figures demonstrating the uncertainties as a function of sky-location:

	1. Our choice of signal population: we consider multiple masses for neutron stars, including one system with an asymmetric mass ratio, and employ a physically uninformative mass prior (that is, flat in component masses).  As Nissanke et al. is primarily interested in recovery of the binary distance, inclination, and sky locations, their work (and all previous MCMC results in the literature) only consider neutron stars of 1.4 solar masses, and employs an unphysical flat prior on the chirp mass (which yields a non-flat prior in component masses).

We have made this distinction clear in the text (see the new paragraph at the end of section 1, and the second paragraph in the conclusion, and the end of section 2.0).

	2. Our choice of signal strength: The high-SNR selection ensures that our reported credible regions are accurate, as it is only in the high-SNR regime that the zero-noise as average posterior assumption becomes valid.  This serves the goal of the study to provide quotable lower-limit estimates on the average parameter recoveries.  This was stated in the last paragraph of section 2.3, and we have modified the text (first paragraph of the conclusion and the new, 5th paragraph of the introduction) to make this goal sufficiently explicit.

Finally, the reviewer points out that the caption for Figure 9 is confusing.  We agree, and have clarified the text (specifically mentioning that the variance and mean of multiple zero-noise results are accurate to different orders in 1/SNR)

### 2013-12-12 doc/paper.tex
Reviewer responses before 12/9 meeting

Within the next five years, it is expected that the Advanced LIGO/Virgo network will have
reached a sensitivity sufficient to enable the routine detection of
gravitational waves.  Beyond the initial detection, the scientific
promise of these instruments relies on the effectiveness of our physical
parameter estimation capabilities. The majority of this effort has been towards the
detection and characterization of gravitational waves from compact
binary coalescence, e.g. the coalescence of binary neutron stars.
While several previous studies have investigated the accuracy of parameter
estimation with advanced detectors, the majority have relied
on approximation techniques such as the Fisher Matrix.  Here we report the
statistical uncertainties that will be achievable for optimal detection candidates
 ($\text{SNR}=20$) using
the full parameter estimation machinery developed by the LIGO/Virgo
Collaboration via Markov-Chain Monte Carlo methods. 
 We find the recovery of the individual masses to be
fractionally within 9\% (15\%) at the 68\% (95\%) credible intervals for equal-mass
systems, and within 1.9\% (3.7\%) for unequal-mass systems.  We also find that the Advanced
LIGO/Virgo network will constrain the locations of binary neutron star mergers to a median
 uncertainty of $5.1$ $\mathrm{deg}^2$ (13.5 $\mathrm{deg}^2$) on the sky.  This region is improved to  2.3 $\mathrm{deg}^2$ $(6$ $\mathrm{deg}^2)$ with the addition of the proposed LIGO India detector to the network.  We also report the
average uncertainties on the luminosity distances and
orbital inclinations of ideal detection candidates that can be
achieved by different network configurations.
% \will{I'm not sure that
  %we want to say this exactly this way---the sources in this study are
 % not randomly distributed in SNR, but rather cluster at rather loud
  %signals.}
\end{abstract}

@@

By employing the full parameter-estimation machinery that will
eventually be used in the Advanced LIGO/Virgo era, our results
 give realistic estimates of the capabilities of
advanced ground-based detectors to characterize BNS systems.  Until
recently, the majority of studies have employed the Fisher-matrix
formalism which was first adapted for gravitational-wave parameter
estimation by \citet{FinnDetection}.  While each of these studies
\citep{PoissonWill,CutlerFlanagan,ArunPE} have pointed out the
limitations and flaws of the Fisher Information Matrix, there has been
relatively few studies which investigate the BNS parameter estimation capabilities
of Advanced LIGO/Virgo using the techniques that will eventually be
employed.  The recent work of \cite{Vallisneri} and our own work
\citep{Inadequacies} has demonstrated that the Fisher matrix cannot
even be treated as a lower bound on the standard deviations of certain
parameters (particularly the masses) measurable in gravitational-wave
detectors.  While previous studies have employed MCMC techniques for specific
parameters,  \citep{Nissanke2011,Nissanke2013,Veitch2012}, none have considered 
the full range of parameters measurable for an optimal detection candidates.

In this paper, we perform a systematic study of the statistical
uncertainties with which the Advanced LIGO/Virgo network will be able
to measure the basic parameters of BNS mergers.  In Section
\ref{PEsection}, we describe the machinery of our parameter estimation
code, \texttt{LALInference}, and its associated MCMC sampler,
\texttt{lalinference\_mcmc}, as well as the frequency-domain
gravitational-wave template we employ.  In Section
\ref{resultsSection}, we qualitatively analyze the posterior
probability density functions for BNS systems with different masses
and extrinsic parameters.  We select three equal-mass and one unequal-mass
 binary systems as prototypical examples of BNS systems.  Each system is analyzed 40 
 times with isotropically selected sky locations and orbital orientations, and with a distance
 such that each signal was injected with a network signal-to-noise ratio of 20.
  The results are divided into three subsections of interest: the recovery of the mass
parameters (Section \ref{massSection}), the recovery of the orbital
inclination and luminosity distance (Section \ref{idSection}), and the
localization of sources on the sky (Section \ref{skySection}).
Finally we provide quantitative 1-dimensional credible intervals on
each parameter in Section \ref{ciSection}.  These results, contained
in Tables \ref{ciTableIntrinsic} and \ref{ciTableExtrinsic}, are
intended as a reference for the optimal BNS parameter estimation
capabilities of Advanced LIGO/Virgo.  We assume that $G=c=1$
throughout.

@@

We are interested in the posterior $p(\thpara | s)$ as it encodes
information about the prior state of knowledge about the problem in
addition to the likelihood of the signal.  For instance, since it is
assumed that compact binaries form homogeneously in the local universe
observable by advanced detectors, we can safely assume that the
population of sources will be distributed homogeneously in volume, i.e.\
with a density proportional to the square of the luminosity distance $D^2$.  
Our prior information can come from mathematical limits in the parameter space 
or from our \textit{a priori} knowledge of astrophysical systems.  
We employ priors which are
\begin{itemize}
\item uniform in component masses from $0.8M_{\odot} \leq M_{1,2} \leq
  30M_{\odot}$, with a minimum chirp mass (see Eq.~\eqref{eqRatioCM})
  of 0.6 $M_{\odot}$,
\item uniform in volume, which implies a luminosity distance prior of
  $p(D)dD \propto D^2 dD$,
\item uniform in coalescence time over the segment of data being
  analyzed,
  \item a delta function at zero spin, and
\item isotropic in all angles
\end{itemize}
The mass priors listed here are the generic priors used in follow-up
analysis, not specifically for BNS systems.  While we could have used
a prior with a lower component mass cutoff for the present study, in
practice the higher prior boundary does not affect our current
analysis.

@@

For a multi-detector network, the noise-weighted inner products
\eqref{innerProduct} combine linearly, allowing us to use the above
formalism with minimal modification.  We integrate the inner product
from a lower-frequency cutoff of $20\text{Hz}$ to the
innermost-stable-circular orbit of the systems, which for
a non-spinning binary is a function only of the total mass:

@@

Finally, note that the PSD that defines the inner
product \eqref{innerProduct} is simply the time-averaged sensitivity
of a given detector to a specific frequency.  Ignoring non-Gaussian
glitches, any stretch of data should contain a specific noise
realization drawn from a Gaussian colored by the PSD.  Without a
realistic model including the actual instrument sensitivity, non-stationary
 noise, and uncertainty in the PSD we cannot accurately
simulate gravitational-wave data to the point of making precise predictions about
parameter estimation capabilities.  Furthermore, the posterior
distribution from a future detection will be influenced by the
realization of the noise at the time of the detection.  Because of
this, we elect to focus only on the statistical errors due to the
flexibility of the waveforms and the expected sensitivity of the
advanced era detectors and suppress any effects of detector noise by
picking $n(t) \equiv 0$.

Consequently, what we recover with the MCMC could be interpreted as
the posterior distribution function for each source averaged over all
possible (Gaussian) noise realizations, assuming we have the correct
PSD and that the signal is sufficiently powerful for the linearized-signal
approximation to be valid.  By selecting a source population with an
 artificial distribution in SNR, we ensure that the
     ``zero-noise'' as average uncertainty approximation is valid.
       This approach has been employed in previous studies to similar effect
        (e.g. \cite{Nissanke2010} and \cite{Sampson2013}).
 To check that our reported statistical errors are not in gross
disagreement with those obtained in the presence of noise, we selected
a particular simulated signal to perform parameter estimation upon in
several noise realizations.  The comparison of the noise-dependent and
noise-free results can be found in Section \ref{noiseSection}.

@@

The difference in uncertainties between the two studies is illustrative.  \cite{Nissanke2013}
considered a realistic distribution of sources motivated by a local galaxy catalogue for
 sources where $D\leq 200\text{Mpc}$, and homogeneously distributed in co-moving
 volume for sources where $D > 200\text{Mpc}$.  These were then limited to those signals
 where $\text{SNR}_{network} \geq 8.5$ (the threshold for detection candidates in the
 current LIGO/Virgo matched-filtering framework) for the MCMC parameter estimation.
   The current study considers a source
 population isotropically distributed on the sky, but at distances such that each signal
 represents an optimal case for parameter estimation.   Fig.\ \ref{compSky} compares
 the sky-area uncertainties that will be available to those considering ``good 
 gravitational-wave events'' (e.g. precision source localization), and the 
 uncertainties that one can gather by considering all gravitational-wave candidates with 
 $\text{SNR}_{network}\geq 8.5$ (e.g. electromagnetic follow-up).  As expected, both source
 populations show a dramatic decrease in parameter estimation uncertainties with the
 addition of the LIGO India site to the advanced detector network.  This holds true even when
 considering the increased number of nominal sources available to the HLVI network 
 configuration in the Nissanke population, where the expansion of the network detection 
 horizon due to LIGO India increases the number of sources to 130, versus 90 in 
 the HLV configuration.  This trend is similar to that observed in \cite{Veitch2012}, contrasting
 the originally proposed four-detector configuration (with two co-located Hanford detectors) to 
 the HLVI configuration.

@@

\begin{figure}[ht!]
  \centering \includegraphics[trim=0cm 0cm 0cm 0cm,
    clip=true,scale=0.52]{noisePDF.pdf}
 \caption{The effects of a non-zero noise realization on the recovered
   PDF for five different Gaussian realizations of the Advanced LIGO
   noise curve.  Each blue-dashed PDF represents the recovery of the same
   $1M_{\odot}/2.5M_{\odot}$ signal in a different noise curve, picked
   at random from the Gaussian colored noise defined by the Advanced
   LIGO power spectral density, while the gray-shaded curve is the
   zero-noise PDF.  Notice how each curve is a Gaussian PDF with different mean and slightly
   different width
   , to the zero-noise PDF.  This
   only holds true for glitch-free data, which is an unrealistic
   idealization when compared to real data.}
 \label{noisePDFs}
\end{figure}

@@

  Of course, this also relies on the assumption that the detector noise is Gaussian.  
  This is the second reason for employing the zero-noise approximation: it is not feasible
to predict the characteristics of realistic (i.e. non-Gaussian)
advanced detector noise.  Attempting to do so would have little chance
of being accurate while requiring a huge computational burden to
combine the results of many MCMC runs in many different noise realizations.
These non-Gaussian excursions will be one of the largest hurdles for parameter estimation
in the advanced detector era.
  Techniques which build
off of the theoretical progress made
in~\cite{Allen:2002jw,Rover:2011qd,Littenberg:2010gf} for including
glitches in the model for the data, and therefore relaxing the
assumptions about stationary and Gaussian noise, are currently under
investigation.

@@

Given the previous reliance of parameter estimation studies on the Fisher
Information Matrix, it is informative to compare the credible intervals quoted here 
to previously reported values in the literature.  We recompute the credible intervals 
of the chirp mass and the symmetric mass ratio ($\eta \equiv m_1 m_2 / M_{tot}^2$)
at the 68\% level (1-$\sigma$) for the $1.4M_{\odot}/1.4M_{\odot}$ system presented 
here.  We then compare the findings of the MCMC to the Fisher matrix predictions using
an identical network configuration (HLV), noise curve, and waveform model.  
We compute the Fisher matrix uncertainties 
using a code previously described in \cite{Inadequacies}.  We find that the MCMC credible
 intervals vary significantly from
the estimates of the Fisher matrix, with the Fisher matrix underestimating the 
fractional uncertainties by as much as a factor of three.  See Table~\ref{FIMvsMCMC}.  This 
result agrees with the conventional wisdom surrounding the Fisher Matrix accuracy,
as well as results where the Fisher Matrix is computed to higher order \citep{Vitale2010}
   This disagreement, while 
   unsurprising, emphasizes the importance of employing the full parameter estimation
    machinery of an MCMC when making physically relevant claims.

\begin{table*}[t!]
\centering
\caption{Median 68\% and 95\% credible intervals for intrinsic parameters for
  each of the four systems considered.  We report the credible
  intervals of quantities measured, as well as the component masses
  and total mass.  Although the results for the HLV and HLVI
  configurations are quantitatively identical, we report them
  separately for consistency.}
  \tabcolsep=0.11cm
      {\renewcommand{\arraystretch}{1.3} 
\begin{tabular}{lcccccccccccccccccccc}

@@

\begin{table*}[t!]
\centering
\caption{Median 68\% and 95\% credible intervals of extrinsic parameters for
  each of the four systems considered.  As expected, there exists a
  substantial improvement in the sky localization capabilities of the
  four-detector HLVI configuration over the three-detector HLV
  configuration.  Note that the solid-angle sky-location credible
  intervals, $\Delta\Omega$, are calculated directly on the 2D sphere,
  not by combining the $\alpha$ and $\delta$ uncertainties.}

@@

\begin{table}[b!]
\centering
\caption{68\% Credible intervals versus 1-$\sigma$ Fisher matrix uncertainties for mass parameters in each system in the HLV configuration.}
  \tabcolsep=0.11cm
    {\renewcommand{\arraystretch}{1.3} 
\begin{tabular}{lcccccccc}

@@

$1M_{\odot}-1M_{\odot}$ & \vline &  0.00497\% & \vline & 0.00205\% & \vline & 0.645\% & \vline & 0.321\% \\
$1.4M_{\odot}-1.4M_{\odot}$ & \vline &  0.00890\% & \vline & 0.00384\% & \vline & 0.745\% & \vline & 0.380\% \\
$1M_{\odot}-2.5M_{\odot}$& \vline &  0.0176\% & \vline & 0.00457\% & \vline & 1.475\% & \vline & 0.400\% \\
$2.5M_{\odot}-2.5M_{\odot}$& \vline &  0.0245\% & \vline & 0.0114\% & \vline & 1.000\% & \vline & 0.612\% \\

@@

In this paper, we performed an MCMC parameter estimation analysis on
the recoverability of basic information about binary neutron stars,
using two projected versions of the Advanced LIGO/Virgo network.  We
focused on the recovery of the two masses, the luminosity distance,
orbital inclination, and the sky location, as these are the six basic
parameters of physical interest to the problem.  We found that,
neglecting the effects of spin, the component masses can be
constrained to within 9\% (15\%) of their true value to a credible level of
68\% (95\%).  This value drops below 1.9\% (3.7\%) for systems with an asymmetric mass
ratio.  These results were summarized in Table \ref{ciTableIntrinsic}.
It should be remembered that these signals were injected at high, 
but not unrealistic, SNRs.  Thus, the results quoted can be thought of as the
average uncertainties for a ``good'' Advanced LIGO/Virgo detection.

@@

Secondly, as stated above, we have performed our study under idealized
detector-noise conditions.  In practice the noise levels of Advanced
LIGO and Advanced Virgo can vary with time and contain occasional
excursions which are highly non-Gaussian.  Unfortunately, there is no
reliable way to predict the sort of non-Gaussian detector glitches and
instrumentation effects that will arise in any advanced
gravitational-wave detector.  The realization of noise will be a
major factor in the deflection of signal PDFs from the idealized cases
presented here.

@@

>>>>

Within the next five years, it is expected that the Advanced LIGO/Virgo network will have
reached a sensitivity sufficient to enable the routine detection of
gravitational waves.  Beyond the initial detection, the scientific
promise of these instruments relies on the effectiveness of our physical
parameter estimation capabilities. The majority of this effort has been towards the
detection and characterization of gravitational waves from compact
binary coalescence, e.g. the coalescence of binary neutron stars.
While several previous studies have investigated the accuracy of parameter
estimation with advanced detectors, the majority have relied
on approximation techniques such as the Fisher Matrix\carl{, while ignoring the rich, non-Gaussian
nature of the gravitational-wave parameter space}.  Here we report the
statistical uncertainties that will be achievable for optimal detection candidates
 ($\text{SNR}=20$) using
the full parameter estimation machinery developed by the LIGO/Virgo
Collaboration via Markov-Chain Monte Carlo methods\carl{, with the goal of supplanting the previously quoted Fisher Matrix bounds}. 
 We find the recovery of the individual masses to be
fractionally within 9\% (15\%) at the 68\% (95\%) credible intervals for equal-mass
systems, and within 1.9\% (3.7\%) for unequal-mass systems.  We also find that the Advanced
LIGO/Virgo network will constrain the locations of binary neutron star mergers to a median
 uncertainty of $5.1$ $\mathrm{deg}^2$ (13.5 $\mathrm{deg}^2$) on the sky.  This region is improved to  2.3 $\mathrm{deg}^2$ $(6$ $\mathrm{deg}^2)$ with the addition of the proposed LIGO India detector to the network.  We also report the
average uncertainties on the luminosity distances and
orbital inclinations of ideal detection candidates that can be
achieved by different network configurations.
\end{abstract}

@@

By employing the full parameter-estimation machinery that will
eventually be used in the Advanced LIGO/Virgo era, our results
 give \sout{realistic} \carl{lower-bound} estimates of the capabilities of
advanced ground-based detectors to characterize BNS systems.  
The majority of studies have employed the Fisher-matrix
formalism which was first adapted for gravitational-wave parameter
estimation by \citet{FinnDetection}.  While each of these studies
\citep{PoissonWill,CutlerFlanagan,ArunPE} have pointed out the
limitations and flaws of the Fisher Information Matrix, there has been
relatively few studies which investigate the BNS parameter estimation capabilities
of Advanced LIGO/Virgo using the techniques that will eventually be
employed.  \sout{The recent work of} \cite{Vallisneri} \sout{and our own work}
\citep{Inadequacies} \sout{ has demonstrated that the Fisher matrix cannot
even be treated as a lower bound on the standard deviations of certain
parameters (particularly the masses) measurable in gravitational-wave
detectors.}

\carl{The Fisher-matrix estimates which the current study seeks to supplant ignore 
the complicated structure of the gravitational-wave parameter space.  The recovered
parameter distributions are often highly non-Gaussian, with multiple peaks and correlations, details
which violate the explicit assumptions of the Fisher-matrix formalism.   By relaxing those constraints, this study
provides a full exploration of the parameter space in the context of optimal detection candidates.  This picture is
necessary to understand the }

In this paper, we perform a systematic study of the statistical
uncertainties with which the Advanced LIGO/Virgo network will be able
to measure the basic parameters of BNS mergers.  In Section
\ref{PEsection}, we describe the machinery of our parameter estimation
code, \texttt{LALInference}, and its associated MCMC sampler,
\texttt{lalinference\_mcmc}, as well as the frequency-domain
gravitational-wave template we employ.  In Section
\ref{resultsSection}, we qualitatively analyze the posterior
probability density functions for BNS systems with different masses
and extrinsic parameters.  We select three equal-mass and one unequal-mass
 binary systems as prototypical examples of BNS systems.  Each system is analyzed 40 
 times with isotropically selected sky locations and orbital orientations, and with a distance
 such that each signal was injected with a network signal-to-noise ratio of 20, \carl{a loudness selected to
 ensure the averages }
  The results are divided into three subsections of interest: the recovery of the mass
parameters (Section \ref{massSection}), the recovery of the orbital
inclination and luminosity distance (Section \ref{idSection}), and the
localization of sources on the sky (Section \ref{skySection}).
Finally we provide quantitative 1-dimensional credible intervals on
each parameter in Section \ref{ciSection}.  \sout{These results, contained
in Tables \ref{ciTableIntrinsic} and \ref{ciTableExtrinsic}, are
intended as a reference for the optimal BNS parameter estimation
capabilities of Advanced LIGO/Virgo.}

\carl{The goal of the current study is to provide an lower-limit ``rule-of-thumb'' for parameter estimation uncertainties 
in Advanced LIGO/Virgo which acknowledges the complexity of the parameter space, and separates this from
 the difficulties introduced by the detector noise for a single event.  These results are collected and summarized in 
 Tables \ref{ciTableIntrinsic} and \ref{ciTableExtrinsic}.  While previous studies have employed 
 MCMC techniques for specific parameters,  \citep{Nissanke2011,Nissanke2013,Veitch2012}, none have considered 
the full range of parameters measurable for an optimal detection candidates, nor have they considered neutron stars
 of masses other than $1.4M_{\odot}$.  We explore
 the recovery of mass parameters for BNS systems with multiple neutron star masses and mass ratios.  
 We assume that $G=c=1$.}

@@

We are interested in the posterior $p(\thpara | s)$ as it encodes
information about the prior state of knowledge about the problem in
addition to the likelihood of the signal.  For instance, since it is
assumed that compact binaries form homogeneously in the local universe
observable by advanced detectors, we can safely assume that the
population of sources will be distributed homogeneously in volume, i.e.\
with a density proportional to the square of the luminosity distance $D^2$.  
Our prior information can come from mathematical limits in the parameter space 
or from our \textit{a priori} knowledge of astrophysical systems.  
We employ priors which are
\begin{itemize}
\item uniform in component masses from $0.8M_{\odot} \leq M_{1,2} \leq
  30M_{\odot}$, with a minimum chirp mass (see Eq.~\eqref{eqRatioCM})
  of 0.6 $M_{\odot}$ 
\item uniform in volume, which implies a luminosity distance prior of
  $p(D)dD \propto D^2 dD$,
\item uniform in coalescence time over the segment of data being
  analyzed,
  \item a delta function at zero spin, and
\item isotropic in all angles
\end{itemize}
The mass priors listed here are the generic priors used in follow-up
analysis, not specifically for BNS systems.  While we could have used
a prior with a lower component mass cutoff for the present study, in
practice the higher prior boundary does not affect our current
analysis.  \carl{Note that the choice of a flat prior in component masses is distinct from \cite{Nissanke2013},
 which employed a prior flat in chirp mass and reduced mass-ratio.}

@@

For a multi-detector network, the noise-weighted inner products
\eqref{innerProduct} combine linearly \carl{so long as the noise is
uncorrelated between detectors}, allowing us to use the above
formalism with minimal modification.  We integrate the inner product
from a lower-frequency cutoff of $20\text{Hz}$ to the
innermost-stable-circular orbit of the systems, which for
a non-spinning binary is a function only of the total mass:

@@

Finally, note that the PSD that defines the inner
product \eqref{innerProduct} is simply the time-averaged sensitivity
of a given detector to a specific frequency.  Ignoring non-Gaussian
glitches, any stretch of data should contain a specific noise
realization drawn from a Gaussian colored by the PSD.  Without a
realistic model including the actual instrument sensitivity, non-stationary
 noise, and uncertainty in the PSD we cannot accurately
simulate gravitational-wave data to the point of making precise predictions about
parameter estimation capabilities.  Furthermore, the posterior
distribution from a future detection will be influenced by the
realization of the noise at the time of the detection.  Because of
this, we elect to focus only on the statistical errors due to the
flexibility of the waveforms and the expected sensitivity of the
advanced era detectors and suppress any effects of detector noise by
picking $n(t) \equiv 0$.  Consequently, what we recover with the MCMC could be interpreted as
the posterior distribution function for each source averaged over all
possible (Gaussian) noise realizations, assuming we have the correct
PSD and that the signal is sufficiently powerful for the linearized-signal
approximation to be valid.  By selecting a source population with an
 artificial distribution in SNR, we ensure that the
     ``zero-noise'' as average uncertainty approximation is valid.

     Previous studies have both employed and studied in detail the consequences of this approach.
   \carl{  In particular, \cite{Nissanke2010} demonstrated that the $n(t)=0$ posterior is equal to the geometric mean over 
     specific noise realizations of the posterior, while \cite{Sampson2013} performed a numerical experiment to 
     demonstrate the validity of the assumption.  We perform a similar numerical experiment in Section \ref{noiseSection}, in which we demonstrate that the uncertainties from a $n(t)=0$ posterior are equivalent to the frequentist
     average over uncertainties from multiple noise realizations. }

@@

The difference in uncertainties between the two studies is illustrative.  \cite{Nissanke2013}
considered a realistic distribution of sources motivated by a local galaxy catalogue for
 sources where $D\leq 200\text{Mpc}$, and homogeneously distributed in co-moving
 volume for sources where $D > 200\text{Mpc}$.  These were then limited to those signals
 where $\text{SNR}_{network} \geq 8.5$ (the \carl{rule-of-thumb} threshold for detection candidates in the
 current LIGO/Virgo matched-filtering framework) for the MCMC parameter estimation.
   The current study considers a source
 population isotropically distributed on the sky, but at distances such that each signal
 represents an optimal case for parameter estimation.   Fig.\ \ref{compSky} compares
 the sky-area uncertainties that will be available to those considering ``good 
 gravitational-wave events'' (e.g. precision source localization), and the 
 uncertainties that one can gather by considering all gravitational-wave candidates with 
 $\text{SNR}_{network}\geq 8.5$ (e.g. electromagnetic follow-up).  As expected, both source
 populations show a dramatic decrease in parameter estimation uncertainties with the
 addition of the LIGO India site to the advanced detector network.  This holds true even when
 considering the increased number of nominal sources available to the HLVI network 
 configuration in the Nissanke population, where the expansion of the network detection 
 horizon due to LIGO India increases the number of sources to 130, versus 90 in 
 the HLV configuration.  This trend is similar to that observed in \cite{Veitch2012}, contrasting
 the originally proposed four-detector configuration (with two co-located Hanford detectors) to 
 the HLVI configuration.

@@

\begin{figure}[ht!]
  \centering \includegraphics[trim=0cm 0cm 0cm 0cm,
    clip=true,scale=0.52]{noisePDF.pdf}
 \caption{The effects of a non-zero noise realization on the recovered
   PDF for five different Gaussian realizations of the Advanced LIGO
   noise curve.  Each blue-dashed PDF represents the recovery of the same
   $1M_{\odot}/2.5M_{\odot}$ signal in a different noise curve, picked
   at random from the Gaussian colored noise defined by the Advanced
   LIGO power spectral density, while the gray-shaded curve is the
   zero-noise PDF.  \carl{Notice how each curve is a Gaussian PDF with different mean but similar 
   width from the zero-noise PDF.  This is to be expected, as the zero-noise mean is identical to the 
   average to first order in $1/\text{SNR}$, while the zero-noise standard deviation is identical to the average 
   standard-deviation
    up to $1/\text{SNR}^3$.  }
   This only holds true for glitch-free data, which is an unrealistic
   idealization when compared to real data.}
 \label{noisePDFs}
\end{figure}

@@

  Of course, this also relies on the assumption that the detector noise is Gaussian.  
  This is the second reason for employing the zero-noise approximation: it is not feasible
to predict the characteristics of realistic (i.e. non-Gaussian)
advanced detector noise.  
Attempting to do so would have little chance
\sout{of being accurate while requiring a huge computational burden to
combine the results of many MCMC runs in many different noise realizations.
These non-Gaussian excursions will be one of the largest hurdles for parameter estimation
in the advanced detector era} \carl{However, as the current study is focused on the lower-bound average uncertainties,
 and not the systematics of a single event (for which we must await a true detection), this assumption does not impact 
 the validity of our results}
Techniques which build
off of the theoretical progress made
in~\cite{Allen:2002jw,Rover:2011qd,Littenberg:2010gf} for including
glitches in the model for the data, and therefore relaxing the
assumptions about stationary and Gaussian noise, are currently under
investigation.

@@

Given the previous reliance of parameter estimation studies on the Fisher
Information Matrix, it is informative to compare the credible intervals quoted here 
to previously reported values in the literature.  We recompute the credible intervals 
of the chirp mass and the symmetric mass ratio ($\eta \equiv m_1 m_2 / M_{tot}^2$)
at the 68\% level \carl{NO, FIX ME}(1-$\sigma$) for the $1.4M_{\odot}/1.4M_{\odot}$ system presented 
here.  We then compare the findings of the MCMC to the Fisher matrix predictions using
an identical network configuration (HLV), noise curve, and waveform model.  
We compute the Fisher matrix uncertainties 
using a code previously described in \cite{Inadequacies}.  We find that the MCMC credible
 intervals \carl{can} vary significantly from
the estimates of the Fisher matrix, with the Fisher matrix underestimating the 
fractional uncertainties by as much as a factor of \sout{three} \carl{two}.  See Table~\ref{FIMvsMCMC}.  This 
result agrees with the conventional wisdom surrounding the Fisher Matrix accuracy,
as well as results where the Fisher Matrix is computed to higher order \citep{Vitale2010}
   This disagreement, while 
   unsurprising, emphasizes the importance of employing the full parameter estimation
    machinery of an MCMC when making physically relevant claims.

\begin{table*}[h!]
\centering
\caption{Median 68\% and 95\% credible intervals for intrinsic parameters for
  each of the four systems considered.  We report the credible
  intervals of quantities measured, as well as the component masses
  and total mass.  Although the results for the HLV and HLVI
  configurations are quantitatively identical, we report them
  separately for consistency.}
  \tabcolsep=0.11cm
      {\renewcommand{\arraystretch}{1.3} 
\begin{tabular}{lcccccccccccccccccccc}

@@

\begin{table*}[h!]
\centering
\caption{Median 68\% and 95\% credible intervals of extrinsic parameters for
  each of the four systems considered.  As expected, there exists a
  substantial improvement in the sky localization capabilities of the
  four-detector HLVI configuration over the three-detector HLV
  configuration.  Note that the solid-angle sky-location credible
  intervals, $\Delta\Omega$, are calculated directly on the 2D sphere,
  not by combining the $\alpha$ and $\delta$ uncertainties.}

@@

\begin{table}[h!]
\centering
\caption{68\% Credible intervals versus \sout{1-$\sigma$} \carl{$2\sigma$} Fisher matrix uncertainties for mass parameters in each system in the HLV configuration.}
  \tabcolsep=0.11cm
    {\renewcommand{\arraystretch}{1.3} 
\begin{tabular}{lcccccccc}

@@

$1M_{\odot}-1M_{\odot}$ & \vline &  0.00497\% & \vline & 0.00396\% & \vline & 0.645\% & \vline & 0.704\% \\
$1.4M_{\odot}-1.4M_{\odot}$ & \vline &  0.00890\% & \vline & 0.00743\% & \vline & 0.745\% & \vline & 0.866\% \\
$1M_{\odot}-2.5M_{\odot}$& \vline &  0.0176\% & \vline & 0.00880\% & \vline & 1.475\% & \vline & 0.862\% \\
$2.5M_{\odot}-2.5M_{\odot}$& \vline &  0.0245\% & \vline & 0.0221\% & \vline & 1.000\% & \vline & 1.299\% \\

@@

In this paper, we performed an MCMC parameter estimation analysis on
the recoverability of basic information about binary neutron stars,
using two projected versions of the Advanced LIGO/Virgo network.  We
focused on the recovery of the two masses, the luminosity distance,
orbital inclination, and the sky location, as these are the six basic
parameters of physical interest to the problem.  It should be remembered that these signals were injected at high, 
but not unrealistic, SNRs \carl{in order to create a reference similar to those found in previous
Fisher Matrix studies, and to validate the ``$n(t) = 0$ as average'' assumption.

 Unlike the Fisher-matrix studies, an MCMC allows us to relax the flawed assumptions of a Gaussian and uni-modal 
 distribtion, and to
fully explore the relevant parameter space.  Given the highly non-Gaussian nature of the explored 
parameter space (see Figures \ref{9dPDF}, \ref{1414masses}, \ref{metaMassPDFs}-\ref{2525SkyLocHLVI} ), this exploration is critical to a complete 
understanding of the gravitational-wave BNS parameter estimation problem}.  
The results quoted can be thought of as the
average \carl{lower-bound} uncertainties for an \carl{optimal} \sout{``good''} Advanced LIGO/Virgo detection.

\carl{For the mass parameters,} we found that, neglecting the effects of spin, the component masses can be
constrained to within 9\% (15\%) of their true value to a credible level of
68\% (95\%) \carl{for systems with equal-mass neutron stars.
}  This value drops below 1.9\% (3.7\%) for systems with an asymmetric masses.
  \carl{It was also found that the fractional uncertainties for equal-mass binary systems are similar at equal SNR.
Only when the masses are unequal are the fractional errors effected.}
These results were summarized in Table \ref{ciTableIntrinsic}.  \carl{The current study is the first 
to quantify neutron star mass measurements in Advanced LIGO/VIRGO  via an MCMC for neutron star masses
other than $1.4M_{\odot}$.}

@@

Secondly, as stated above, we have performed our study under idealized
detector-noise conditions.  In practice the noise levels of Advanced
LIGO and Advanced Virgo can vary with time and contain occasional
excursions which are highly non-Gaussian.  Unfortunately, there is no
reliable way to predict the sort of non-Gaussian detector glitches and
instrumentation effects that will arise in any advanced
gravitational-wave detector.  The realization of noise will be a
major factor in the deflection of signal PDFs from the idealized, \carl{lower-bound} cases
presented here.

@@

\pagebreak
